% ===========================================================================
\section{Arithmetization summary}\label{sec:summary}
% ===========================================================================

We summarize the dimensions and constraint counts of the SHA-256
UAIR$^+$ above.

\begin{center}
\small
\begin{tabular}{ll}
\toprule
\textbf{Quantity} & \textbf{Value} \\
\midrule
Number of compressions $N$        & $7$ \\
Rows per compression              & $68$ \quad ($4$ init-prefix $+\ 64$ round-update outputs) \\
Active rows                       & $7 \cdot 68 + 4 = 480$ \\
Trace length $n = 2^{\nu}$        & $2^{9} = 512$ \quad (smallest power of $2$ $\ge 480$) \\
Prime tuple $\mathbf{q}$          & $()$ \quad (single $\Q[X]$ branch) \\
\midrule
Binary-polynomial columns         & $20$ \quad ($9$ public $+\ 11$ witness) \\
\quad of which booleanity-confined to $\bitpoly{32}$ & all $20$ (free) \\
\quad of which committed witness  & $11$ \\
Integer columns                   & $9$ \quad ($9$ public $+\ 0$ witness) \\
\midrule
Total committed columns           & $\mathbf{11}$ \\
Total public-input columns        & $\mathbf{18}$ \quad ($9$ binary-poly $+\ 9$ int) \\
\midrule
Constraint families (excluding lookups) & $12$ \quad (C\ref{con:c1}, C\ref{con:c2}, C\ref{con:c4}, C\ref{con:c6}, C\ref{con:c7}--C\ref{con:c9}, \\
                                  & \phantom{$12$ \quad } C\ref{con:c10}, C\ref{con:c11}, C\ref{con:ff-a}, C\ref{con:ff-e}, C\ref{con:msg-init}, C\ref{con:mu-zero}) \\
Verifier-side public-column structural checks & \Cref{cons:public-checks}: bitpoly\{32\}, compensator-zero, compensator-zero, etc. \\
Virtual binary-polynomial residuals & $3$ \quad ($r_{\mathrm{ch}_1},\,r_{\mathrm{ch}_2},\,r_{\mathrm{maj}}$) \\
Affine-combination lookups        & $3$ families ($\Ch_{ef}$, $\Ch_{\neg e,g}$, $\Maj$) \\
\midrule
Maximum constraint degree (in trace MLEs) & $1$ \quad (excluding selector-gated zero-ideal pins) \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Why $11$ committed witness columns suffice}\label{sec:why-11}

The Zinc+ paper's idealized arithmetization (Section~8.2.3 there) uses
a witness layout with separate columns for each of the spec's Table~9
materialization rows ($B_1, B_2, B_3$) and a separate integer-typed
carry column for each modular sum. Our deployed arithmetization removes
all of these:
\begin{itemize}
\item The Table~9 materialization columns $B_1, B_2, B_3$ are not
committed; instead, the affine combinations of \eqref{eq:rch1}--\eqref{eq:rmaj}
are declared as virtual binary-polynomial columns and pinned into
$\bitpoly{32}$ per coefficient by the booleanity sumcheck. This saves
$3$ committed columns and removes the corresponding $3$ materialization
identities.
\item The five integer-typed carry columns
$\mu_W,\mu_a,\mu_e,\mu_{ff,a},\mu_{ff,e}$ are packed into a single
binary-polynomial column $\bp{W_{\mu\text{-packed}}}$ via
$\mathrm{ShiftR}$ virtuals (\Cref{rem:mu-budget}). This saves
$5$ committed columns and gives the carry range-checks for free
through booleanity.
\end{itemize}
The cost is a single extra range-pin (C\ref{con:mu-zero}) and a
modestly more involved bit-budget accounting. We retain the
multi-byte $\bp{W_a}, \bp{W_e}, \bp{W_W}, \bp{W_{\Sigma_0}},
\bp{W_{\Sigma_1}}, \bp{W_{\sigma_0}}, \bp{W_{\sigma_1}},
\bp{W_{u_{ef}}}, \bp{W_{u_{\neg e,g}}}, \bp{W_{\Maj}}$ as committed
columns because they participate non-trivially in the rotation,
register-update, and Ch/Maj lookups; they cannot be virtualized
without inflating the constraint degree.

\subsection{Verifier-side public-column inspection}

All $9$ binary-polynomial public columns and all $9$ integer public
columns are part of \texttt{public\_trace}, which the verifier
holds before the IOPP runs. The verifier discharges every
public-column structural property by direct row-wise inspection
(\Cref{cons:public-checks}): bitpoly\{32\}-membership for every
public binary-poly column, zero-on-active-rows for the five
compensators $\col{PA\_C\_*}$, zero-on-inner-rows plus boundary
closed form for the two tail compensators $\col{PA\_R\_*\_COMP}$,
and the $\{0,1\}$ / round-constant / message structure of the
remaining public columns. This costs no algebraic constraint, no
selector column, and no additional commitment; the marginal cost is
$O(n)$ verifier work folded into the existing public-MLE
construction loop.

Several of the public binary-polynomial columns are themselves
functions of other public data and could be supplied by the verifier
through shadow-running the SHA computation:
$\col{PA\_OV\_SIG0/SIG1/LSIG0/LSIG1}$ are determined by
$\col{PA\_A}, \col{PA\_E}$ and the rotation arithmetic; the tail
compensators $\col{PA\_R\_CH2\_COMP}$ and $\col{PA\_R\_MAJ\_COMP}$ are
determined by the last two rows of $\col{PA\_A}$ and $\col{PA\_E}$
through \eqref{eq:compensators}. The arithmetization treats them as
public input rather than recomputed in-circuit; the verifier
enforces the algebraic constraints, while the public values are
computed offline by either the prover or a precompute step.

\subsection{Composition with downstream UAIRs}

The chosen layout exposes:
\begin{itemize}
\item $\col{PA\_A}\rowidx{0..4}$ and $\col{PA\_E}\rowidx{0..4}$, the
input digest $H_0$;
\item $\col{PA\_A}\rowidx{68N..68N+4}$ and $\col{PA\_E}\rowidx{68N..68N+4}$,
the output digest $H_N$;
\item $\col{PA\_M}\rowidx{68i..68i+16}$ for $i \in [0,N)$, the per-block
message words;
\end{itemize}
all as public input cells. A downstream UAIR (e.g.\ the merged
SHA+ECDSA UAIR of the implementation) can therefore read $H_N$
directly from this UAIR's public input without any in-circuit
cross-binding constraint, which is the composition pattern adopted by
the implementation's \texttt{Sha256CompressionSliceUair} +
\texttt{EcdsaUair} side-by-side merge.
