\section{Soundness}
\label{sec:soundness}

We argue informally that \Cref{cons:bitop} preserves the
knowledge-soundness error of the host PIOP up to a negligible additive
term linear in the size of $\Sbit$. The argument is structural: every
new claim added by the construction is bound to existing claims either
algebraically (sumcheck claim-value reconstruction) or by closed-form
substitution (\Cref{eq:bitop-open}).

\subsection{Threat model}

We assume the host PIOP is knowledge-sound when restricted to the
existing classes of trace claims (base columns and row-shift virtual
columns). Let $\epsilon_{\mathrm{host}}$ denote its soundness error.
We argue that the augmented PIOP has soundness error
\[
  \epsilon_{\mathrm{host}} \;+\; m \cdot \epsilon_{\mathrm{mp}} \;+\; \negligible(\lambda),
\]
where $\epsilon_{\mathrm{mp}}$ is the per-source soundness loss of the
multi-point evaluation argument (typically $\nu / q$), and
$m = |\Sbit|$.

\subsection{Failure modes}

A malicious prover $\mathcal{P}^*$ can attempt to cheat in at most
three places:

\paragraph{(F1) The base column's lifted opening
$\lmle{v_{j_\ell}}(r_0)$ is wrong.} The verifier reads the lifted
opening to form the open evals \cref{eq:base-open} and substitutes
them into the mp\_eval subclaim \cref{eq:final-subclaim}. If
$\lmle{v_{j_\ell}}(r_0) \ne \mle{v_{j_\ell}}(r_0)$, then with
probability $\ge 1 - \negligible(\lambda)$ over $\alpha$ and $r_0$,
$\psia(\lmle{v_{j_\ell}}(r_0)) \ne \mle{\psia(v_{j_\ell})}(r_0)$, and
\cref{eq:final-subclaim} fails. This is a Schwartz–Zippel argument on
the polynomial that interpolates the committed projected MLE,
identical to the host PIOP and unchanged by our construction.

\paragraph{(F2) The bit-op claim at $r^{**}$ is wrong.} Suppose
$b_{\ell}^{**}$ disagrees with $M_{\ell}(r^{**})$ for some $\ell$.
Conditioned on the sumcheck claim-value equation passing, this implies
the constraint polynomial $C$ evaluates to a value at $(r^*, r^{**})$
that is consistent with the prover's answers but not with any honest
trace; standard sumcheck soundness pins this event to probability
$\le \nu \deg(C)/q$. If instead the prover compensates by tampering
$\mathrm{down}_{\mathrm{bit, eval}}$ in mp\_eval consistently with
$b_{\ell}^{**}$, the residual claim $e_{\ell}$ at $r_0$ will be
inconsistent with the equivariance identity, caught by (F3).

\paragraph{(F3) The bit-op residual claim at $r_0$ is wrong.}
Conditioned on (F1) not occurring, the verifier holds the correct
$\lmle{v_{j_\ell}}(r_0)$, and \cref{eq:bitop-open} yields the correct
\[
  M_{\ell}(r_0) \;=\; \psia\!\big(\bitop_{\ell}(\lmle{v_{j_\ell}}(r_0))\big)
\]
deterministically — \emph{the prover has no freedom over $\tilde e_\ell$.}
The mp\_eval subclaim \cref{eq:final-subclaim} therefore reduces to a
single algebraic identity whose left- and right-hand sides match if
and only if every $b_{\ell'}^{**}$ was consistent with
$M_{\ell'}(r^{**})$ (and likewise for the base sources and their
row-shifts). By multi-point-evaluation soundness, an incorrect
$b_{\ell}^{**}$ propagates to a violation of \cref{eq:final-subclaim}
with probability at least $1 - \nu/q$ over $r_0$. We pay this loss
once per extra source, hence $m \cdot \epsilon_{\mathrm{mp}}$ overall.

\subsection{Why no fresh extractor is required}

A standard concern when adding new prover messages is whether
extraction of the witness needs new machinery. Here, the only new
prover messages are the $m$ scalars $\{b_{\ell}^{**}\}$, and they are
\emph{not} commitments — they are values of MLEs that the verifier
can recompute (in the soundness analysis, by virtue of holding the
extracted base columns).

Concretely, a knowledge extractor $\mathcal{E}$ for the augmented
PIOP runs the host extractor unchanged to recover the base columns
$(v_j)_{j \in \Sup}$. It then defines the bit-op virtual columns by
$\bitop_{\ell *}(v_{j_\ell})$ — these are deterministic functions of
the base columns. Soundness of the host PIOP guarantees the base
columns satisfy the host's constraints. By
\Cref{lem:equivariance}, $\mle{\bitop_{\ell *} v_{j_\ell}}$ agrees
with $M_{\ell}$ on the boolean cube, hence as multilinear polynomials.
Therefore the augmented constraint $C$ — which by hypothesis was
satisfied at the sumcheck endpoint — is satisfied row-wise by the
extracted columns. The extractor needs no new randomness, no new
oracle queries, and no new rewinding beyond what the host PIOP
already does.

\subsection{Quantitative summary}

\begin{proposition}[Soundness, informal]
\label{prop:soundness}
Let $\Pi$ denote the host multilinear PIOP and $\Pi^+$ its
augmentation by \Cref{cons:bitop} with bit-op set $\Sbit$ of size $m$.
If $\Pi$ has knowledge-soundness error $\epsilon_{\mathrm{host}}$, then
$\Pi^+$ has knowledge-soundness error
\[
  \epsilon_{\mathrm{host}} \;+\; m \cdot \epsilon_{\mathrm{mp}}.
\]
\end{proposition}

The $m \cdot \epsilon_{\mathrm{mp}}$ term is identical in form to the
soundness loss the host PIOP already pays per row-shift virtual
column.
